The resurgence of the linear optics quantum interferometer --- recent advances \& applications

### 2018-05-30 arxiv/paper.tex
Added a folder for the arxiv version.

%\documentclass[aps,rmp,twocolumn,amsmath,amssymb,nofootinbib,superscripta
%ddress]{revtex4}

\documentclass[times,final]{elsarticle}
\usepackage{revip}
\usepackage{framed,multirow,threeparttable}
\usepackage{hyperref,mathrsfs,amsmath,amssymb,latexsym,array,hhline}
\newcommand{\bra}[1]{\langle#1|}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\op}[2]{\hat{\textbf{#1}}_{#2}}
\newcommand{\dagop}[2]{\hat{\textbf{#1}}_{#2}^\dag}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

@@

\journal{Reviews in Physics}

\begin{document}
\verso{Tan and Rohde}

\begin{frontmatter}

%
% Title
%

\title{The resurgence of the linear optics \sihui{quantum} interferometer --- recent advances \& applications}

%
% Authors
%

\author[1,2]{Si-Hui \snm{Tan}}
\cortext[cor1]{\peter{sihui\_tan@sutd.edu.sg}}

\author[3]{Peter P. \snm{Rohde}\corref{cor1}}

\address[1]{Singapore University of Technology and Design, Singapore 
University of Technology and Design, 8 Somapah Rd, Singapore 487372}
\address[2]{Centre for 
Quantum Technologies, National University of Singapore, Block S15, 3 
Science Drive 2, Singapore 117543, Singapore}
\address[3]{Centre for Quantum Software \& Information (CQSI), Faculty 
of Engineering \& Information Technology, University of Technology 
Sydney, NSW 2007, Australia}

\received{}
\finalform{}
\accepted{}
\availableonline{}
\communicated{}

@@

\begin{keyword}
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
\MSC 41A05\sep 41A10\sep 65D05\sep 65D17
%% Keywords
\KWD Keyword1\sep Keyword2\sep Keyword3
\end{keyword}

\end{frontmatter}

@@

Linear optics play a prominent role in quantum information processing \cite{bib:LovettKok}. Photons make fantastic `flying' qubits, and are readily used for quantum communication \cite{bib:Northup14} and quantum key distribution \cite{bib:Lo14}. In 2001, Knill, Laflamme and Milburn (KLM) showed that efficient quantum computing is possible using only linear optical components, that is single photons, beamsplitters, phase shifters and photon counting  \sihui{for spatially encoded qubits \cite{bib:KLM01}. At about the same time, Koashi, Yamamoto, and Imoto (KYI) essentially came to the same result for polarization-encoded qubits \cite{Koashi01}.} Furthermore, an optical implementation of quantum protocols can potentially piggyback on existing communication infrastructure. Combined with the relative ease of transporting quantum states of light, as compared to say trapped ions or superconducting qubits, linear optics forms a powerful platform for quantum information processing.

The promises offered by linear optics have spearheaded many \sihui{scientific and} technical advancements. One of the major successes for linear optics is \sihui{the detection of gravitational waves by the LIGO experiments based on the Michelson interferometer \cite{bib:LIGO2015}. The further application of squeezed light to the detectors promises to reduce quantum noise, which currently limits their performance \cite{bib:Barsotti18}. Another advancement is} the paradigm shift from bulk optics to integrated photonics \cite{bib:Meany15}. The advantage of using such integrated photonics over bulk optics is that it is more stable against phase fluctuations, and miniaturized. This has increased the scalability of optical implementations of quantum information protocols. Furthermore, it is now possible to have single-photon sources and detectors together with linear-optical networks on a silica chip \cite{bib:Sprengers11, bib:Silverstone14}. Having all components on chip reduces coupling losses which would be crucial for fault-tolerance \cite{bib:Li15}. \sihui{Experimental advancements and limitations of photonic implementations of quantum information processing are discussed at length in another review \cite{bib:Flamini2018}.}

In the last decade, single photons have been produced mainly through \sihui{nonlinear optical processes know as spontaneous parametric down-conversion (SPDC) and spontaneous four-wave mixing (SFWM) \cite{bib:Eisaman11,bib:Xiong16,bib:Solntsev17}.} \sihui{Both SPDC and SFWM produce} single photons in correlated pairs that has been a staple source of entanglement. \sihui{However, SPDC uses up one pump photons per pair, while SFWM requires two.} Through developments in production, the number of high quality indistinguishable photons that can be simultaneously produced have increased steadily over the years. Single-photon pairs, once a novelty, are now run-of-the-mill business for quantum optical \sihui{laboratories}, and three or four photons have become commonplace \cite{bib:Spagnolo12, bib:Tillmann13, bib:Spring13,bib:Broome13, bib:Crespi13}. Experiments using up to ten photons have been demonstrated \cite{bib:WangChen16, bib:Chen17}. We have also seen the meteoric rise of quantum dots in the role for producing single photons \cite{bib:Ding16,bib:WangHe16, bib:213601}. They are a scalable source of single photons which are highly indistinguishable, and are produced on demand--a significant advantage over \sihui{nonlinear optical processes which are inherently probabilistic.}

Another approach to linear optics is to pack more features known as 
degrees of freedom onto a single photon, and to increase the degree 
of control over them. It is possible to manipulate degrees of
freedom like polarization, time-of-arrival, and orbital angular momentum \cite{bib:Tillmann2015,bib:Bozinovic2013, 
bib:Nicolas2014, bib:Humphreys2013, bib:Donohue2013}. Since KLM \sihui{and KYI}, new proposals 
for linear optical quantum computations (LOQC) have surfaced 
\cite{bib:Nielsen04,bib:BrowneRudolph05,bib:Gimeno-Segovia15,bib:Pant17}. 
\sihui {Some of them support new functionalities like secure delegated quantum 
computations \cite{bib:Barz12,bib:Fisher14, bib:Tan16}}. Others do not promise universality in their processing, but still performs tasks such as sampling \cite{bib:Aaronson11}, and \sihui{encrypted quantum walks \cite{bib:Rohde_qw12}} that might be hard or 
impossible with a classical system.  Such methods open up new ways to engineer 
quantum states and quantum operations, and expand the toolbox of baseline 
resources we have for implementing quantum computations, which could prove 
crucial in the race to achieve quantum supremacy.

We begin in Section \ref{sec:math} with an overview of the mathematical treatment of linear optical transformation on single photons, and in Section \ref{sec:encoding} with a description of some commonly used qubit encodings. Any unitary transformation acting on the spatial label of single photons can be efficiently decomposed into a network of beamsplitters and phase-shifters. We discuss this, and an extension to handle additional internal states, in Section \ref{sec:decomp}. For applying any linear optical circuit, one needs an accurate description of its unitary representation. A full quantum tomography is costly and time-consuming. Some practical methods have been devised to do this reconstruction without resorting to tomography, and these are described in Section \ref{sec:reconstruction} \sihui{while quantum tomography is briefly discussed in Section \ref{secQPT}}. In Section \ref{sec:state}, we recap the experimental breakthroughs in state preparation of single photons, and some entangled quantum states. This is followed by advances in time-bin based architecture for linear optical networks, \sihui{photodectors, and optical switches in Sections \ref{sec:timebins}, \ref{sec:detector}, and \ref{sec:switch} respectively}. Last, but not least, we review applications for linear optical interferometry in \sihui{Section \ref{sec:applications}}.

@@

Unlike polarisation or dual-rail encoding, time-bin encoding does not lend itself to `native' single-qubit operations. Rather, fast switching can be used to spatially separate neighbouring time-bins, implement a beamsplitter operation between them, before converting back to time-bin encoding. \sihui{Typical switch times that can be achieved are in the picoseconds range \cite{bib:Dietrich16}.} An experiment has built on this idea to implement a two-qubit gate by fast switching to a polarization basis \cite{bib:Humphreys2013}. Another promising advance is an ultrafast measurement technique for time bins based on converting into frequency bins \cite{bib:Donohue2013}. Scalable networks using time-bin encoding are also possible using a loop-based architecture \cite{bib:Motes14}. This is discussed in more detail in Section \ref{sec:timebins}.

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.55\columnwidth]{reck}
\caption{Efficient decomposition of arbitrary linear optics 
networks into a sequence of beamsplitters and phase-shifters. 
\sihui{This setup describes a generalized $n\times n$ multi-port Mach-Zehnder interferometer. Adapted from \cite{bib:Reck1994}.}} \label{fig:Reck}	
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\columnwidth]{configMZI}
\caption{ \sihui{(a) A decomposition of arbitrary linear optics networks into a sequence of subunitaries, $D_i$, made up of Mach-Zehnder interferometers, and is a universal linear-optical processing unit. These subunitaries are controlled via their photon amplitudes ($\alpha_{i,j}$) and phases ($\phi_{i,j}$). (b) A state-of-art implementation of a universal linear-optical processing unit is shown. The processor has 6 inputs entering from the left, and 6 outputs leaving to the right, and is constructed from a cascade of 15 Mach-Zehnder interferometers. The classical control and processing unit sets the parameters for the interferometer, and actively cools it. Adapted from \cite{bib:Carolan15}.}} \label{fig:Carolan}	
\end{figure}

Alternatively, \sihui{Mach-Zehnder} interferometers can also be employed as building blocks instead of beamsplitters and phase-shifters \cite{bib:Englert2001}. \sihui{Such a decomposition and its implementation on an integrated photonic chip are shown in Fig.~\ref{fig:Carolan}. } Later, it was shown that any nontrivial beamsplitter, that does more than permuting modes or adding phases to them, is universal for linear optics \cite{bib:Bouland2014}. However, they do not provide an explicit construction for arbitrary unitaries.

If the linear optical transformations \sihui{is to} be realized on various degrees of freedom of light, then it is possible to realize a $n\times n$ arbitrary unitary transformation, where $n=n_s n_p$ for $n_s$ spatial modes, and $n_p$ internal modes, by a sequence of $\mathcal{O}(n_s^2 n_p)$ beamsplitters and $\mathcal{O}(n_s^2)$ internal transformations \cite{bib:Dhand2015}. This approach reduces the required number of beamsplitters but increases the total number of optical elements needed by a factor of 2.

@@

\peter{When demonstrating quantum protocols experimentally it is generally absolutely critical to characterise the device prior to actual protocol execution. Tomography can be used to reconstruct the full density operator of an output state, and quantum process tomography to reconstruct the full process matrix implemented by a device. While various other measurement schemes can partially reconstruct quantum states or processes, or simply reconstruct certain parameters quantifying them, tomography is complete -- fully and uniquely reconstructing state and process matrices, from which any other measure can be determined. Importantly, state and process tomography require no interference or multi-qubit gates -- just separable single-qubit measurements in different bases. But unfortunately they require a number of measurements growing exponentially with the size of the system.

State tomography is achieved by measuring expectation values in a complete basis for the state Hilbert space. For one qubit, for example, a complete basis is the Pauli basis, which wields the completeness relation,}
\begin{align}
\hat\rho = \text{tr}({\hat\rho}) + \text{tr}(\hat{X}\hat{\rho})\hat{X} + \text{tr}({\hat{Y}\hat\rho})\hat{Y} + \text{tr}(\hat{Z}{\hat\rho})\hat{Z}.
\end{align}
\peter{These traces correspond exactly to the expectation values in the respective bases, which are easily determined using separable measurements. \sihui{Sometimes, it is possible to be able to do a tomography from an incomplete number of measurements, {\it i.e.} a number that is less than the number of expectation values, and it is possible using compressed sensing provided some sparsity conditions are met \cite{bib:GLFBE2010,bib:Titchener16,bib:Oren16}. In cases where one only wishes to know if a given state is close to a presumed state, one can perform a certification test instead of state tomography \cite{bib:Aolita15}.}

Process tomography employs state tomography as a primitive, applying an equivalent completeness relation, but across the space of quantum processes, which may be decomposed into a basis input to output states. For a single qubit process this may be represented as,}
\begin{align} \label{eq:process_matrix}
\mathcal{E}(\hat\rho) = \sum_{i,j=1}^4 \chi_{i,j} \hat{\sigma}_i\hat\rho\,\hat{\sigma}_j^\dag,
\end{align}
\peter{where $\hat\sigma_i$ are the Pauli operators. This generalises to an arbitrary number of qubits. Here $\chi$ is the process matrix in the Pauli basis, which fully characterises the quantum process.}
%\cite{bib:Mitchell03,bib:Obrien04,bib:Lobino08,bib:Saleh11}.

@@

The photonic states that are most commonly employed in linear optics protocols can be divided into Fock states (e.g single-photon), and entangled states, such as EPR, GHZ and cluster states. We will consider advances in each of these. \sihui{Quantum state engineering using quantum interferometers has also attracted some interest, and is discussed at the end of this section.}

@@

Solid-state single photon sources are versatile and efficient sources of single photons, however, the photons they produce have suffered from the lack of indistinguishability that is necessary for typical quantum information processing applications. Recent developments using resonant excitation of quantum dots were able to overcome these limitations. Laser pulses are used to excite the electronic resonance of the quantum dots and trigger the emission of high quality single photons \cite{bib:WangHe16}
%\cite{bib:Ding16,bib:213601,bib:Ates09, bib:Dirk10,bib:
%1748,bib:Jayakumar13,bib:Wei14,bib:Muller14,bib:Unsleber15,
%bib:237403,bib:Sweeney14,bib:WangHe16}.

@@

A closely related state is the NOON state, written as
\begin{align}
\ket{\psi^{\rm (NOON)}_n}=\frac{1}{\sqrt{2}}(\ket{n,0}+\ket{0,n})\ ,
\end{align}
in the second quantization representation, \sihui{and the basis is the same as that for eq.~(\ref{eq:dualrail})}. This state is structurally similar to an EPR state, but its large photon-number give it great metrological power in quantum phase estimation, allowing saturation of the Heisenberg limit -- a physical bound on accuracy in phase estimation.

@@

\begin{figure}
\centering
\begin{subfigure}{.5\columnwidth}
  \centering
  \includegraphics[width=.5\columnwidth]{cluster_state}
  \caption{}
  \label{fig:cluster}
\end{subfigure}%
\begin{subfigure}{.5\columnwidth}
  \centering
  \includegraphics[width=.7\columnwidth]{fusion}
  \caption{}
  \label{fig:fusion}
\end{subfigure}
\caption{(a) The representation of cluster states as a graph. Vertices represent qubits initialised into the $\ket{+}$ state, while edges represent the application on {\sc cphase} gates, the ordering of which is irrelevant. (b) Fusion gates for joining smaller cluster states into larger ones. Both require only a single polarising beamsplitter, and waveplates. The type-I and -II fusion gates consume 1 or 2 qubits (photons) respectively, creating an edge between the remaining graphs. The type-I gate consumes one fewer photon, but requires number-resolved detection on the detected mode. The type-II gate requires only on/off photodetection, but consumes an additional photon. Thus, the type-I gate can be employed to fuse two Bell pairs into a 3-qubit cluster state, whereas the type-II gate will only grow clusters when beginning with at least 3 qubits per cluster. }
\label{fig:test}
\end{figure}

@@

\sihui{
\subsection{Quantum state engineering based on the multi-port Mach-Zehnder interferometer}
Linear optical inteferometers have been used for quantum state engineering. One well-known example is that of optical state truncation, sometimes also referred to as quantum scissors, and it happens by way of a process known as projection synthesis.  Projection synthesis works by projecting an input state onto an output state by putting the former through a multiport interferometer, and then postselecting on a certain photon number output. Optical state truncation is a use of projection synthesis to use a continuous variable state represented in the Fock-state basis, $\ket{\phi}=\sum_{n=0}^\infty c_n\ket{n}$, to produce the normalized form of $\sum_{n=0}^d c_n\ket{n}$. An example of the interferometer used for optical state truncation for the case $d=1$ is shown in Fig.~\ref{fig:qscissors}. The idea of projection synthesis was first conceived for the special case of $d=2$ \cite{bib:Pegg98, bib:Barnett99}, and later extended to $d=3$ \cite{bib:Koniorczyk2000}, and arbitrarily large $d$ \cite{bib:Miranowicz07, bib:Goyal13}. Notably, the scheme proposed in \cite{bib:Goyal13} uses multimode single photons and repeated uses of $d=2$ quantum scissors to extend the depth of truncation. 
}

@@

\sihui{When} using time-bin encoding of optical qubits, the obvious question is\sihui{: how do we perform single-qubit and two-qubit} operations upon them. In \cite{bib:Motes14}, a dual-loop architecture was presented for implementing arbitrary passive linear-optics on photonic pulse-trains with time-bin separation $\tau$, shown in Fig.~\ref{fig:FL_arch}. The inner loop has length exactly $\tau$, while the outer one has length $>n\tau$ ($n$ is the 
number of optical modes). The architecture is controlled via three 
dynamically controlled beamsplitters. The first and last need only be on/off switches, whose sole purpose is to couple in the prepared pulse-train, 
keep it within the outer loop for the required duration, and then couple 
out of the outer loop, yielding the transformed pulse-train. The central 
beamsplitter must be able to implement arbitrary classically-controlled 
beamsplitter operations.

@@

The above description applies to passive linear optics, which is sufficient for protocols such as {\sc BosonSampling}, but insufficient for universal optical quantum computation, which requires the addition of ancillary states, and measurement with fast-feedforward. To address this, it was shown in \cite{Rohde15} that \sihui{partial measurements can be implemented by coupling in and out arbitrary subsets of the optical modes. The coupling is in turn achieved via dynamically preparing ancillary pulse-trains (using the already existing source) that are classically controlled by time-resolved measurements at the output, and changing the switching sequence conditional on these measurement outcomes.}

@@

\peter{
Some of the protocols thus far require switching between different optical modes. \sihui{One way to achieve this experimentally for time-bin qubits uses} a Mach-Zehnder interferometer with a dynamic, classically-controlled phase-shift in one path \cite{bib:Takesue14} (see Fig.~\ref{figMZ}). The variable phase may be implemented in a number of ways using various types of phase modulators, such as electro-optic or acousto-optic modulators. With this construction setting the phase to $\phi=0$ performs the identity operation, whereas for $\phi=\pi$ it performs a swap gate. \sihui{Ref.~\cite{bib:Takesue14} reports a switching time of 100 picoseconds for this type of switch.}
}

\begin{figure}[!htb]
\centering
\includegraphics[width=0.55\columnwidth]{MZ}	
\caption{} \label{figMZ}
\end{figure}

>>>>

\documentclass[pra,singlecolumn,superscriptaddress,twoside,pra]{revtex4}

%\documentclass[times,final]{elsarticle}
%\usepackage{revip}
\usepackage{framed,multirow,threeparttable}
\usepackage{hyperref,mathrsfs,amsmath,amssymb,latexsym,array,hhline}
\newcommand{\bra}[1]{\langle#1|}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\op}[2]{\hat{\textbf{#1}}_{#2}}
\newcommand{\dagop}[2]{\hat{\textbf{#1}}_{#2}^\dag}
\usepackage{graphicx}
\usepackage{subcaption}
\captionsetup{compatibility=false}

@@

\begin{document}
%
% Title
%

\title{The resurgence of the linear optics quantum interferometer --- recent advances \& applications}

%
% Authors
%
\author{Si-Hui Tan}
\affiliation{Singapore University of Technology and Design, 20 Dover Drive, Singapore 138682}
\affiliation{Centre for Quantum Technologies, National University of Singapore, Block S15, 3 Science Drive 2, Singapore 117543, Singapore}
%
\author{Peter P.~Rohde}
\affiliation{Centre for Quantum Software \& Information (CQSI), Faculty 
of Engineering \& Information Technology, University of Technology 
Sydney, NSW 2007, Australia}

@@

\date{\today}
%\pacs{03.67.Ac,03.67.Dd,03.67.Lx,05.40.Fb}
\maketitle

@@

Linear optics play a prominent role in quantum information processing \cite{bib:LovettKok}. Photons make fantastic `flying' qubits, and are readily used for quantum communication \cite{bib:Northup14} and quantum key distribution \cite{bib:Lo14}. In 2001, Knill, Laflamme and Milburn (KLM) showed that efficient quantum computing is possible using only linear optical components, that is single photons, beamsplitters, phase shifters and photon counting for spatially encoded qubits \cite{bib:KLM01}. At about the same time, Koashi, Yamamoto, and Imoto (KYI) essentially came to the same result for polarization-encoded qubits \cite{Koashi01}. Furthermore, an optical implementation of quantum protocols can potentially piggyback on existing communication infrastructure. Combined with the relative ease of transporting quantum states of light, as compared to say trapped ions or superconducting qubits, linear optics forms a powerful platform for quantum information processing.

The promises offered by linear optics have spearheaded many scientific and technical advancements. One of the major successes for linear optics is the detection of gravitational waves by the LIGO experiments based on the Michelson interferometer \cite{bib:LIGO2015}. The further application of squeezed light to the detectors promises to reduce quantum noise, which currently limits their performance \cite{bib:Barsotti18}. Another advancement is the paradigm shift from bulk optics to integrated photonics \cite{bib:Meany15}. The advantage of using such integrated photonics over bulk optics is that it is more stable against phase fluctuations, and miniaturized. This has increased the scalability of optical implementations of quantum information protocols. Furthermore, it is now possible to have single-photon sources and detectors together with linear-optical networks on a silica chip \cite{bib:Sprengers11, bib:Silverstone14}. Having all components on chip reduces coupling losses which would be crucial for fault-tolerance \cite{bib:Li15}. Experimental advancements and limitations of photonic implementations of quantum information processing are discussed at length in another review \cite{bib:Flamini2018}.

In the last decade, single photons have been produced mainly through nonlinear optical processes know as spontaneous parametric down-conversion (SPDC) and spontaneous four-wave mixing (SFWM) \cite{bib:Eisaman11,bib:Xiong16,bib:Solntsev17}. Both SPDC and SFWM produce single photons in correlated pairs that has been a staple source of entanglement. However, SPDC uses up one pump photons per pair, while SFWM requires two. Through developments in production, the number of high quality indistinguishable photons that can be simultaneously produced have increased steadily over the years. Single-photon pairs, once a novelty, are now run-of-the-mill business for quantum optical laboratories, and three or four photons have become commonplace \cite{bib:Spagnolo12, bib:Tillmann13, bib:Spring13,bib:Broome13, bib:Crespi13}. Experiments using up to ten photons have been demonstrated \cite{bib:WangChen16, bib:Chen17}. We have also seen the meteoric rise of quantum dots in the role for producing single photons \cite{bib:Ding16,bib:WangHe16, bib:213601}. They are a scalable source of single photons which are highly indistinguishable, and are produced on demand--a significant advantage over nonlinear optical processes which are inherently probabilistic.

Another approach to linear optics is to pack more features known as 
degrees of freedom onto a single photon, and to increase the degree 
of control over them. It is possible to manipulate degrees of
freedom like polarization, time-of-arrival, and orbital angular momentum \cite{bib:Tillmann2015,bib:Bozinovic2013, 
bib:Nicolas2014, bib:Humphreys2013, bib:Donohue2013}. Since KLM and KYI, new proposals 
for linear optical quantum computations (LOQC) have surfaced 
\cite{bib:Nielsen04,bib:BrowneRudolph05,bib:Gimeno-Segovia15,bib:Pant17}. 
Some of them support new functionalities like secure delegated quantum 
computations \cite{bib:Barz12,bib:Fisher14, bib:Tan16}. Others do not promise universality in their processing, but still performs tasks such as sampling \cite{bib:Aaronson11}, and encrypted quantum walks \cite{bib:Rohde_qw12} that might be hard or 
impossible with a classical system.  Such methods open up new ways to engineer 
quantum states and quantum operations, and expand the toolbox of baseline 
resources we have for implementing quantum computations, which could prove 
crucial in the race to achieve quantum supremacy.

We begin in Section \ref{sec:math} with an overview of the mathematical treatment of linear optical transformation on single photons, and in Section \ref{sec:encoding} with a description of some commonly used qubit encodings. Any unitary transformation acting on the spatial label of single photons can be efficiently decomposed into a network of beamsplitters and phase-shifters. We discuss this, and an extension to handle additional internal states, in Section \ref{sec:decomp}. For applying any linear optical circuit, one needs an accurate description of its unitary representation. A full quantum tomography is costly and time-consuming. Some practical methods have been devised to do this reconstruction without resorting to tomography, and these are described in Section \ref{sec:reconstruction} while quantum tomography is briefly discussed in Section \ref{secQPT}. In Section \ref{sec:state}, we recap the experimental breakthroughs in state preparation of single photons, and some entangled quantum states. This is followed by advances in time-bin based architecture for linear optical networks, photodectors, and optical switches in Sections \ref{sec:timebins}, \ref{sec:detector}, and \ref{sec:switch} respectively. Last, but not least, we review applications for linear optical interferometry in Section \ref{sec:applications}.

@@

Unlike polarisation or dual-rail encoding, time-bin encoding does not lend itself to `native' single-qubit operations. Rather, fast switching can be used to spatially separate neighbouring time-bins, implement a beamsplitter operation between them, before converting back to time-bin encoding. Typical switch times that can be achieved are in the picoseconds range \cite{bib:Dietrich16}. An experiment has built on this idea to implement a two-qubit gate by fast switching to a polarization basis \cite{bib:Humphreys2013}. Another promising advance is an ultrafast measurement technique for time bins based on converting into frequency bins \cite{bib:Donohue2013}. Scalable networks using time-bin encoding are also possible using a loop-based architecture \cite{bib:Motes14}. This is discussed in more detail in Section \ref{sec:timebins}.

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.4\columnwidth]{reck}
\caption{Efficient decomposition of arbitrary linear optics 
networks into a sequence of beamsplitters and phase-shifters. This setup describes a generalized $n\times n$ multi-port Mach-Zehnder interferometer. Adapted from \cite{bib:Reck1994}.} \label{fig:Reck}	
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=0.8\columnwidth]{configMZI}
\caption{(a) A decomposition of arbitrary linear optics networks into a sequence of subunitaries, $D_i$, made up of Mach-Zehnder interferometers, and is a universal linear-optical processing unit. These subunitaries are controlled via their photon amplitudes ($\alpha_{i,j}$) and phases ($\phi_{i,j}$). (b) A state-of-art implementation of a universal linear-optical processing unit is shown. The processor has 6 inputs entering from the left, and 6 outputs leaving to the right, and is constructed from a cascade of 15 Mach-Zehnder interferometers. The classical control and processing unit sets the parameters for the interferometer, and actively cools it. Adapted from \cite{bib:Carolan15}.} \label{fig:Carolan}	
\end{figure}

Alternatively, Mach-Zehnder interferometers can also be employed as building blocks instead of beamsplitters and phase-shifters \cite{bib:Englert2001}. Such a decomposition and its implementation on an integrated photonic chip are shown in Fig.~\ref{fig:Carolan}. Later, it was shown that any nontrivial beamsplitter, that does more than permuting modes or adding phases to them, is universal for linear optics \cite{bib:Bouland2014}. However, they do not provide an explicit construction for arbitrary unitaries.

If the linear optical transformations is to be realized on various degrees of freedom of light, then it is possible to realize a $n\times n$ arbitrary unitary transformation, where $n=n_s n_p$ for $n_s$ spatial modes, and $n_p$ internal modes, by a sequence of $\mathcal{O}(n_s^2 n_p)$ beamsplitters and $\mathcal{O}(n_s^2)$ internal transformations \cite{bib:Dhand2015}. This approach reduces the required number of beamsplitters but increases the total number of optical elements needed by a factor of 2.

@@

When demonstrating quantum protocols experimentally it is generally absolutely critical to characterise the device prior to actual protocol execution. Tomography can be used to reconstruct the full density operator of an output state, and quantum process tomography to reconstruct the full process matrix implemented by a device. While various other measurement schemes can partially reconstruct quantum states or processes, or simply reconstruct certain parameters quantifying them, tomography is complete -- fully and uniquely reconstructing state and process matrices, from which any other measure can be determined. Importantly, state and process tomography require no interference or multi-qubit gates -- just separable single-qubit measurements in different bases. But unfortunately they require a number of measurements growing exponentially with the size of the system.

State tomography is achieved by measuring expectation values in a complete basis for the state Hilbert space. For one qubit, for example, a complete basis is the Pauli basis, which wields the completeness relatio,n
\begin{align}
\hat\rho = \text{tr}({\hat\rho}) + \text{tr}(\hat{X}\hat{\rho})\hat{X} + \text{tr}({\hat{Y}\hat\rho})\hat{Y} + \text{tr}(\hat{Z}{\hat\rho})\hat{Z}.
\end{align}
These traces correspond exactly to the expectation values in the respective bases, which are easily determined using separable measurements. Sometimes, it is possible to be able to do a tomography from an incomplete number of measurements, {\it i.e.} a number that is less than the number of expectation values, and it is possible using compressed sensing provided some sparsity conditions are met \cite{bib:GLFBE2010,bib:Titchener16,bib:Oren16}. In cases where one only wishes to know if a given state is close to a presumed state, one can perform a certification test instead of state tomography \cite{bib:Aolita15}.

Process tomography employs state tomography as a primitive, applying an equivalent completeness relation, but across the space of quantum processes, which may be decomposed into a basis input to output states. For a single qubit process this may be represented as,
\begin{align} \label{eq:process_matrix}
\mathcal{E}(\hat\rho) = \sum_{i,j=1}^4 \chi_{i,j} \hat{\sigma}_i\hat\rho\,\hat{\sigma}_j^\dag,
\end{align}
where $\hat\sigma_i$ are the Pauli operators. This generalises to an arbitrary number of qubits. Here $\chi$ is the process matrix in the Pauli basis, which fully characterises the quantum process.
%\cite{bib:Mitchell03,bib:Obrien04,bib:Lobino08,bib:Saleh11}.

@@

The photonic states that are most commonly employed in linear optics protocols can be divided into Fock states (e.g single-photon), and entangled states, such as EPR, GHZ and cluster states. We will consider advances in each of these. Quantum state engineering using quantum interferometers has also attracted some interest, and is discussed at the end of this section.

@@

Solid-state single photon sources are versatile and efficient sources of single photons, however, the photons they produce have suffered from the lack of indistinguishability that is necessary for typical quantum information processing applications. Recent developments using resonant excitation of quantum dots were able to overcome these limitations. Laser pulses are used to excite the electronic resonance of the quantum dots and trigger the emission of high quality single photons \cite{bib:WangHe16}.
%\cite{bib:Ding16,bib:213601,bib:Ates09, bib:Dirk10,bib:
%1748,bib:Jayakumar13,bib:Wei14,bib:Muller14,bib:Unsleber15,
%bib:237403,bib:Sweeney14,bib:WangHe16}.

@@

A closely related state is the NOON state, written as
\begin{align}
\ket{\psi^{\rm (NOON)}_n}=\frac{1}{\sqrt{2}}(\ket{n,0}+\ket{0,n})\ ,
\end{align}
in the second quantization representation, and the basis is the same as that for eq.~(\ref{eq:dualrail}). This state is structurally similar to an EPR state, but its large photon-number give it great metrological power in quantum phase estimation, allowing saturation of the Heisenberg limit -- a physical bound on accuracy in phase estimation.

@@

\begin{figure}
\centering
\begin{subfigure}{.5\columnwidth}
  \centering
  \includegraphics[width=.5\columnwidth]{cluster_state}
  \caption{}
  \label{fig:cluster}
\end{subfigure}%
\begin{subfigure}{.5\columnwidth}
  \centering
  \includegraphics[width=.65\columnwidth]{fusion}
  \caption{}
  \label{fig:fusion}
\end{subfigure}
\caption{(a) The representation of cluster states as a graph. Vertices represent qubits initialised into the $\ket{+}$ state, while edges represent the application on {\sc cphase} gates, the ordering of which is irrelevant. (b) Fusion gates for joining smaller cluster states into larger ones. Both require only a single polarising beamsplitter, and waveplates. The type-I and -II fusion gates consume 1 or 2 qubits (photons) respectively, creating an edge between the remaining graphs. The type-I gate consumes one fewer photon, but requires number-resolved detection on the detected mode. The type-II gate requires only on/off photodetection, but consumes an additional photon. Thus, the type-I gate can be employed to fuse two Bell pairs into a 3-qubit cluster state, whereas the type-II gate will only grow clusters when beginning with at least 3 qubits per cluster. }
\label{fig:test}
\end{figure}

@@

\subsection{Quantum state engineering based on the multi-port Mach-Zehnder interferometer}
Linear optical inteferometers have been used for quantum state engineering. One well-known example is that of optical state truncation, sometimes also referred to as quantum scissors, and it happens by way of a process known as projection synthesis.  Projection synthesis works by projecting an input state onto an output state by putting the former through a multiport interferometer, and then postselecting on a certain photon number output. Optical state truncation is a use of projection synthesis to use a continuous variable state represented in the Fock-state basis, $\ket{\phi}=\sum_{n=0}^\infty c_n\ket{n}$, to produce the normalized form of $\sum_{n=0}^d c_n\ket{n}$. An example of the interferometer used for optical state truncation for the case $d=1$ is shown in Fig.~\ref{fig:qscissors}. The idea of projection synthesis was first conceived for the special case of $d=2$ \cite{bib:Pegg98, bib:Barnett99}, and later extended to $d=3$ \cite{bib:Koniorczyk2000}, and arbitrarily large $d$ \cite{bib:Miranowicz07, bib:Goyal13}. Notably, the scheme proposed in \cite{bib:Goyal13} uses multimode single photons and repeated uses of $d=2$ quantum scissors to extend the depth of truncation.

@@

When using time-bin encoding of optical qubits, the obvious question is: how do we perform single-qubit and two-qubit operations upon them. In \cite{bib:Motes14}, a dual-loop architecture was presented for implementing arbitrary passive linear-optics on photonic pulse-trains with time-bin separation $\tau$, shown in Fig.~\ref{fig:FL_arch}. The inner loop has length exactly $\tau$, while the outer one has length $>n\tau$ ($n$ is the 
number of optical modes). The architecture is controlled via three 
dynamically controlled beamsplitters. The first and last need only be on/off switches, whose sole purpose is to couple in the prepared pulse-train, 
keep it within the outer loop for the required duration, and then couple 
out of the outer loop, yielding the transformed pulse-train. The central 
beamsplitter must be able to implement arbitrary classically-controlled 
beamsplitter operations.

@@

The above description applies to passive linear optics, which is sufficient for protocols such as {\sc BosonSampling}, but insufficient for universal optical quantum computation, which requires the addition of ancillary states, and measurement with fast-feedforward. To address this, it was shown in \cite{Rohde15} that partial measurements can be implemented by coupling in and out arbitrary subsets of the optical modes. The coupling is in turn achieved via dynamically preparing ancillary pulse-trains (using the already existing source) that are classically controlled by time-resolved measurements at the output, and changing the switching sequence conditional on these measurement outcomes.

@@

Some of the protocols thus far require switching between different optical modes. One way to achieve this experimentally for time-bin qubits uses a Mach-Zehnder interferometer with a dynamic, classically-controlled phase-shift in one path \cite{bib:Takesue14} (see Fig.~\ref{figMZ}). The variable phase may be implemented in a number of ways using various types of phase modulators, such as electro-optic or acousto-optic modulators. With this construction setting the phase to $\phi=0$ performs the identity operation, whereas for $\phi=\pi$ it performs a swap gate. Ref.~\cite{bib:Takesue14} reports a switching time of 100 picoseconds for this type of switch.

\begin{figure}[!htb]
\centering
\includegraphics[width=0.4\columnwidth]{MZ}	
\caption{An optical switch for time-bin qubits that uses a Mach-Zehnder interferometer with a dynamic, classicallly-controlled phase-shift in one path.}
\label{figMZ}
\end{figure}
